\section*{Abstract}

% \begin{itemize}
%     \item Brief introduction to the topic of landing zones in cloud environments
%     \item Statement of the problem: the concept exists across providers, but differs in definition, structure, and emphasis
%     \item Objective of the report: to compare provider concepts and derive a common definition
%     \item Short indication of the method: qualitative comparative analysis based on provider documentation and selected technical literature
%     \item Summary of the main result: overview of similarities, differences, and a proposed provider-independent definition
% \end{itemize}

Cloud landing zones are widely used to structure enterprise cloud adoption, yet major providers define and implement the concept in different ways. This paper compares landing zone concepts across Microsoft Azure, Amazon Web Services, Google Cloud, Alibaba Cloud, and IBM Cloud to clarify their common purpose and derive a provider-independent definition. The study uses a qualitative, document-based comparative analysis of official provider documentation and technical literature on cloud architecture, multi-cloud environments, and Infrastructure as Code. The comparison focuses on five dimensions: security, governance, scalability, resource organization, and standardization and automation. The analysis shows that all providers treat landing zones as prepared environments that enable controlled workload deployment through security controls, governance mechanisms, resource structures, and repeatable setup processes. However, they differ in terminology, hierarchy models, policy inheritance, automation mechanisms, and the degree to which the concept is tied to provider-specific services or deployable architecture packages. The paper concludes that a landing zone is best understood not as a fixed architecture pattern, but as a standardized, scalable, and largely automated cloud foundation that establishes organizational, security, and technical guardrails for compliant workload deployment and operation. This shared definition supports clearer architectural comparison in multi-cloud and hybrid-cloud contexts.


\vspace{2ex}
\textbf{Keywords:}

Cloud Landing Zone; Multi-Cloud; Cloud Governance; Cloud Architecture; Infrastructure as Code; Provider-Independent Definition